Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
DSN changes

This last point shows that, while there may be a number of unique differentials
for a given collision class, and potentially many collision classes within
some neighborhood of the original collision class, by analyzing the structure
of the input block differential, systems can potentially prohibit successful
collision attacks by preventing structures of input block differences. Note
too how the graph presented in \ref{Fig:Families48} shows a distinct grouping
into three disconnected subgraphs, which fall along input block differential
family boundaries.

@@

- Theory behind Logical Cryptanalysis:
    - Distance heuristic.
    - $\precsim$.
- Tie into Trustworthiness of hash functions.
- Demonstrated metrics.
- Demonstrated 35k new collisions.
- Implications to real world systems.

>>>>

This last point shows that, while there may be a number of unique differentials
for a given collision class, and potentially many collision classes within
some neighborhood of the original collision class, by analyzing the structure
of the input block differential, systems can potentially prohibit successful
collision attacks by preventing structures of input block differences. Note
too how the graph presented in Figure \ref{Fig:Families48} shows a distinct grouping
into three disconnected subgraphs, which fall along input block differential
family boundaries.

@@

We have demonstrated in this paper that our theories for logical
cryptanalysis hold, and lead to a 35 fold increase in collision
discovery when compared to the existing literature.  By employing our
presented distance heuristic and the lattice structure uncovered by
the $\precsim$ relation, we can evaluate the trustworthiness of a hash
function as the inverse of the flexibility of known collisions for
that hash function.  This allows us to measure the structural
weaknesses of a hash function and determine if extant collisions are sufficient
to render a hash function untrustworthy, and likewise to understand if
necessary conditions related to the underlying structure are met to
ensure trustworthiness.

We demonstrate the utility of five metrics of the flexibility present
in a collision on the basis of 1) the number of unique differentials
within a collision class; 2) the size of the unit distance
neighborhood for a collision class; 3) the maximum count of zeros in a
binary representation of a colliding block; 4) the existance of blocks
which collide under multiple initial values; and 5) the presence of
ASCII valued inplut blocks within a collision.  These metrics proved
capable when evaluating the flexibility of collisions in the published
literature, and showed that flexibility is a multi-dimensional
quantity that must be compared over multiple distance metrics, as not all
collisions dominate universally.

Our empirical results furthermore demonstrated that concerns over
industrial use of deprecated hash functions, specifically MD4, are
real, and that structural weaknesses exist which are highly
exploitable and limit the trustworthiness of any system which employs
such hash functions.  We hope the presented results lead to a
reduction in the use of such deprecated hash functions, and that new
hash functions are periodically evaluated when collisions are
discovered to understand if these collisions also indicate the
presence of structural weaknesses.
